### 2014-11-14 TableSummarization/Paper/TableSummarization.tex
split, first pass on formal

%\documentclass{vldb}
\documentclass{sig-alternate}
%\documentclass[12pt,letterpaper]{article}
%\documentclass[12pt]{extreport}
%\documentstyle[amsmath,amsthm,amssymb,twocolumn]{article}
%\newcommand{\mytitle}{Evaluating the Crowd with Confidence}

%\usepackage{amsmath}
\makeatletter
\let\@copyrightspace\relax
\makeatother
\pdfpagewidth=8.5in
\pdfpageheight=11in

\renewcommand{\baselinestretch}{1.0}

\usepackage{enumitem}
\usepackage{times}
\usepackage{subfigure}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,color}
\usepackage{verbatim}
\usepackage{framed}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{framed}
\usepackage[normalem]{ulem}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{example}[definition]{Example}
\newcounter{prob}
\newtheorem{problem}[prob]{Problem}

\newcommand{\agp}[1]{\textcolor{red}{Aditya: #1}}
\newcommand{\agpdel}[1]{\textcolor{blue}{\bf ||Del: #1||}}

\newcommand{\squishlist}{
   \begin{list}{$\bullet$}
    { \setlength{\itemsep}{0pt}
      \setlength{\parsep}{2pt}
      \setlength{\topsep}{2pt}
      \setlength{\partopsep}{0pt}
    }
}
\newcommand{\stitle}[1]{\vspace{0.5em}\noindent\textbf{#1}}
\newcommand{\squishend}{\end{list}}
\newcommand{\eat}[1]{}
\newcommand{\papertext}[1]{#1}
\newcommand{\techreporttext}[1]{}

\newcommand{\calD}{\mathcal{D}\xspace}

\title{Interactive Summarization of Relational Tables}

\author{
Manas Joglekar\\Stanford University\\manasrj@stanford.edu
\and
Hector Garcia-Molina\\Stanford University\\hector@cs.stanford.edu
\and
Aditya Parameswaran\\Stanford University\\adityagp@cs.stanford.edu
}
\begin{document}
\maketitle

\begin{abstract}
In this paper we present {\em smart drill down},
a system for interactively exploring a relational table
to discover ``interesting'' groups of tuples.
Each group of tuples is described by a {\em rule}.
For instance, the rule $(a, b, \star, 1000)$ tells us that
there are a $1000$ tuples with value $a$ in the first column and $b$
in the second column (and any value in the third column).
Smart drill down presents an analyst with a list of rules that
together describe interesting aspects of the table.
The analyst can tailor the definition of interesting,
and can interactively drill down on an existing rule to
explore that part of the table.
We describe an
algorithm for finding the approximately optimal list of rules, and an optimized
sampling scheme for efficiently summarizing large tables. Finally, we
demonstrate the usefulness of smart drill down by performing experiments on
a real dataset.
\end{abstract}

\input{intro}

\input{formal}

\input{algorithms}

\input{exp}

\input{extensions}

\input{related}

% TODO: Experiments on Sampling schemes.

\begin{comment}
Notes based on Explanation Tables paper: Prove that our problem is NP-Hard to solve exactly. Something like the Set Cover Problem.

The way we deal with overlap, is that the key differentiating factor?

We don't have the binary thing. Is it useful? Maybe we can summarize better without it. We are solving a different problem, optimising for coverage rather than explaining the binary attribute. They are optimising for KL divergence of max entropy distribution. They mention many other papers that optimize for cardinality or similar. We do weight times cardinality, which is a compromise between descriptiveness and coverage, and allow user to change the weight function. But should check those papers out anyway.

\end{comment}

{\small 
\bibliographystyle{plain}
\bibliography{TableSummarization}
}

\end{document}