### 2014-11-24 TableSummarization/Paper/TableSummarization.tex
Addressed Aditya's comments. Changes in experiments.

%\documentclass{vldb}
\documentclass{sig-alternate}
%\documentclass[12pt,letterpaper]{article}
%\documentclass[12pt]{extreport}
%\documentstyle[amsmath,amsthm,amssymb,twocolumn]{article}
%\newcommand{\mytitle}{Evaluating the Crowd with Confidence}

\usepackage{amsmath}
\makeatletter
\let\@copyrightspace\relax
\makeatother
\pdfpagewidth=8.5in
\pdfpageheight=11in

\usepackage{enumitem}
\usepackage{times}
\usepackage{subfigure}
\let\proof\relax
\let\endproof\relax
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx,color}
\usepackage{verbatim}
\usepackage{framed}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{framed}
\usepackage[normalem]{ulem}

\renewcommand{\baselinestretch}{1.0}

\renewcommand*\ttdefault{cmvtt}
\usepackage[T1]{fontenc}

\usepackage{floatrow}
\floatsetup[table]{font=scriptsize}

\newtheorem*{theorem*}{Theorem}
\newtheorem{theorem}{Theorem}
\newtheorem*{lemma*}{Lemma}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{example}[definition]{Example}
\newcounter{prob}
\newtheorem{problem}[prob]{Problem}

\newcommand{\agp}[1]{\textcolor{red}{Aditya: #1}}
\newcommand{\mrj}[1]{\textcolor{green}{#1}}
\newcommand{\agpdel}[1]{\textcolor{blue}{\bf ||Del: #1||}}

\newcommand{\squishlist}{
   \begin{list}{$\bullet$}
    { \setlength{\itemsep}{0pt}
      \setlength{\parsep}{2pt}
      \setlength{\topsep}{2pt}
      \setlength{\partopsep}{0pt}
    }
}
\newcommand{\stitle}[1]{\vspace{0.5em}\noindent\textbf{#1}}
\newcommand{\squishend}{\end{list}}
\newcommand{\eat}[1]{}
\newcommand{\papertext}[1]{#1}
\newcommand{\techreporttext}[1]{#1}

\newcommand{\calD}{\mathcal{D}\xspace}

\title{Smart Drill-Down: Interactive Summarization of Relational Tables}

\author{
Manas Joglekar\\Stanford University\\manasrj@stanford.edu
\and
Hector Garcia-Molina\\Stanford University\\hector@cs.stanford.edu
\and
Aditya Parameswaran\\University of Illinois\\adityagp@illinois.edu
}

\begin{document}

\maketitle

\begin{abstract}
In this paper we present {\em smart drill down},
a system for interactively exploring a relational table
to discover ``interesting'' groups of tuples.
Each group of tuples is described by a {\em rule}.
For instance, the rule $(a, b, \star, 1000)$ tells us that
there are a $1000$ tuples with value $a$ in the first column and $b$
in the second column (and any value in the third column).
Smart drill down presents an analyst with a list of rules that
together describe interesting aspects of the table.
The analyst can tailor the definition of interesting,
and can interactively drill down on an existing rule to
explore that part of the table.
We describe an
algorithm for finding the approximately optimal list of rules, and an optimized
sampling scheme for efficiently summarizing large tables. Finally, we
demonstrate the usefulness of smart drill down by performing experiments on
real datasets.
\end{abstract}

\input{intro}

\input{formal}

\input{algorithms}

\input{exp}

\input{extensions}

\input{related}

{\small 
\bibliographystyle{plain}
\bibliography{TableSummarization}
}
\techreporttext{\input{appendix}}

\end{document}