### 2015-02-12 TableSummarization/Paper/conc.tex
Changed image font size, and some other changes

%!TEX root = TableSummarization.tex

\section{Conclusion}\label{sec:conclusion}
We have presented a new data exploration operator called smart drill down.
Like traditional drill down, it allows an analyst
to quickly discover interesting value patterns (rules) that occur frequently
(or that represent high values of some metric attribute)
across diverse parts of a table.

We presented an algorithm for optimally selecting rules
to display, as well as a scheme for performing such selections
based on data samples. Working with samples makes smart drill down
relatively insensitive to the size of the table.

Our experimental results on our experimental prototype show that smart drill down
is fast enough to be interactive under various
realistic scenarios. We also showed that the accuracy is high
when sampling is used, and when the maximum weight ($m_w$)
approximation is used. Moreover, we have a tunable parameter $minSS$ that the user can tweak to tradeoff performance of smart drill down for the accuracy of the rules.